Missing The Point: Non-Convergence in Iterative Imputation Algorithms

### 2020-11-13 Manuscript.Rmd
Combine figures and add labels

---
title: "Missing The Point"
author: "Hanne Oberman"
date: "7-9-2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
bibliography: ref.bib
---

```{r setup, include=FALSE}
# environment
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
load("Data/results_25_75.Rdata")
# load data
load("Data/example.Rdata")
load("Data/example_chains.Rdata")

@@

# Missing The Point: Non-Convergence in Iterative Imputation Algorithms

@@

## Abstract

@@

## Intro

@@

## Identifying non-convergence

@@

## Simulation Set-Up

@@

### Aims

@@

### Data generating mechanism

@@

### Estimands

@@

### Methods

@@

### Performance measures

@@

## Simulation Results

@@

### Diagnostic Methods

<!-- The next four plots show the autocorrelations in each parameter plotted against the number of iterations.  -->

```{r echo=FALSE}
p <- simulation_plot(v="ac.max.mu.Y", name = "Autocorrelation chain means")
p # %>% plotly::ggplotly()
```

Autocorrelation in the chain means decreases rapidly in the first few iterations. The decrease is substantive until $it \geq 6$. This means that there is some initial trending within chains, but the average imputed value quickly reaches stationarity. These results hold irrespective of the missingness condition.

```{r echo=FALSE}
p <- simulation_plot(v="ac.max.sigma.Y", name = "Autocorrelation chain variances")
p # %>% plotly::ggplotly() 
```

Autocorrelation in the chain variances show us something similar. The number of iterations that is required to reach non-improving autocorrelations is somewhat more ambiguous than for chain means, but generally around $it \geq 10$. We do not observe a systematic difference between missingness conditions here either.

```{r echo=FALSE}
p <- simulation_plot(v="ac.max.qhat", name = "Autocorrelation scientific estimate")
p # %>% plotly::ggplotly() 
```

There is more autocorrelation in the scientific estimates than in the univariate parameters (chain means and chain variances). We observe the highest autocorrelations in conditions where 75% of cases are incomplete. Overall, the autocorrelations reach a plateau when $it \geq 20$ to $30$. There is no clear effect of the missingness mechanisms [unnecessary?].

```{r echo=FALSE}
p <- simulation_plot(v="ac.max.lambda", name = "Autocorrelation novel theta")
p # %>% plotly::ggplotly() 
```

The autocorrelation in the novel parameter exhibits a similar trend to the autocorrelation in the scientific estimates. Trending in this parameter diminishes when $it \geq 20$.

<!-- The next four plots show the $\widehat{R}$ in the different parameters. -->

```{r echo=FALSE}
p <- simulation_plot(v="r.hat.max.mu.Y", name = "R hat chain means")
p # %>% plotly::ggplotly() 
```

We observe that $\widehat{R}$-values of the chain means generally decreases as a function of the number of iterations. An exception to this observation is the initial increase when $3 \leq it \leq 5$ [interpret?? due to initialization or is there really more mixing initially??]. After the first couple of iterations, the mixing between chain means generally improves until $it \geq 30$ to $40$. There is no apparent differentiation between the missingness conditions.

```{r echo=FALSE}
p <- simulation_plot(v="r.hat.max.sigma.Y", name = "R hat chain variances")
p # %>% plotly::ggplotly() 
```

The mixing between chain variances mimics the mixing between chain means almost perfectly. Irrespective of the missingness condition, the $\widehat{R}$-values taper off around $it \geq 30$.

```{r echo=FALSE}
p <- simulation_plot(v="r.hat.max.qhat", name = "R hat scientific estimate")
p # %>% plotly::ggplotly() 
```

With the scientific estimate as parameter we observe very similar $\widehat{R}$-values again. We do, however, see some differences between missingness conditions. Conditions where 75% of the cases are incomplete show more extreme non-mixing. The overall trend remains the same: about 30 iterations are required before mixing stops improving substantially.

```{r echo=FALSE}
p <- simulation_plot(v="r.hat.max.lambda", name = "R hat novel theta")
p # %>% plotly::ggplotly() 
```

$\widehat{R}$-values of the novel parameter show a trend similar to the chain means and chain variances. [add something about conditions??]

@@

### Performance Measures

@@

## Discussion

@@

<!-- ## Check if heuristic for AC works -->

@@

## Case Study

>>>>

---
title: "Missing The Point: Non-Convergence in Iterative Imputation Algorithms"
author: "Hanne Oberman"
date: "7-9-2020"
output:
  bookdown::html_document2
bibliography: ref.bib
---

```{r setup, include=FALSE}
# environment
knitr::opts_chunk$set(echo = TRUE, fig.width = 10, fig.width = 10)
library(tidyverse)
library(mice)
library(patchwork)
# load data
load("Data/parameters.Rdata")
load("Data/results_25_75.Rdata")
load("Data/example.Rdata")
load("Data/example_chains.Rdata")

@@

@@

# Abstract

@@

# Intro

@@

# Identifying non-convergence

@@

```{r eval=FALSE, include=FALSE}
# # maybe add ESS instead of AC???
# LaplacesDemon::ESS(parameters[parameters$m == 1 & parameters$sim == 1, "qhat"])
```

# Simulation Set-Up

@@

## Aims

@@

## Data generating mechanism

@@

## Estimands

@@

## Methods

@@

## Performance measures

@@

# Simulation Results

@@

## Diagnostic Methods

Figure \@ref(fig:ac) shows the autocorrelations in the chain means (panel A), chain variances (panel B), regression estimates (panel C), and eigenvalues (panel D).
<!-- The next four plots show the autocorrelations in each parameter plotted against the number of iterations.  -->

```{r ac, fig.cap = "Autocorrelation (AC) with different parameters.", echo=FALSE}
ac_means <- simulation_plot(v="ac.max.mu.Y", name = "AC chain means") 
ac_vars <- simulation_plot(v="ac.max.sigma.Y", name = "AC chain variances")
ac_betas <- simulation_plot(v="ac.max.qhat", name = quote("AC"~beta))
ac_lambdas <- simulation_plot(v="ac.max.lambda", name = quote("AC"~lambda))
(ac_means + ac_vars) / (ac_betas + ac_lambdas) + plot_layout(guides="collect") + plot_annotation(tag_levels = 'A')

Autocorrelation in the chain means decreases rapidly in the first few iterations (see \@ref(fig:ac)A). The decrease is substantive until $it \geq 6$. This means that there is some initial trending within chains, but the average imputed value quickly reaches stationarity. These results hold irrespective of the missingness condition.

Autocorrelation in the chain variances show us something similar (see \@ref(fig:ac)B). The number of iterations that is required to reach non-improving autocorrelations is somewhat more ambiguous than for chain means, but generally around $it \geq 10$. We do not observe a systematic difference between missingness conditions here either.

There is more autocorrelation in the scientific estimates than in the chain means and chain variances (see \@ref(fig:ac)C). We observe the highest autocorrelations in conditions where 75% of cases are incomplete. Overall, the autocorrelations reach a plateau when $it \geq 20$ to $30$. There is no clear effect of the missingness mechanisms [unnecessary?].

The autocorrelation in the eigenvalues exhibits a similar trend to the autocorrelation in the scientific estimates (see \@ref(fig:ac)D). Trending in this parameter diminishes when $it \geq 20$.

Figure \@ref(fig:rh) shows the potential scale reduction factor in the chain means (panel A), chain variances (panel B), regression estimates (panel C), and eigenvalues (panel D).
<!-- The next four plots show the $\widehat{R}$ in the different parameters. -->
```{r rh, fig.cap = "Potential scale reduction factor (Rh) with different parameters.", echo=FALSE}
rh_means <- simulation_plot(v="r.hat.max.mu.Y", name = "Rh chain means") 
rh_vars <- simulation_plot(v="r.hat.max.sigma.Y", name = "Rh chain variances")
rh_betas <- simulation_plot(v="r.hat.max.qhat", name = quote("Rh"~beta))
rh_lambdas <- simulation_plot(v="r.hat.max.lambda", name = quote("Rh"~lambda))
(rh_means + rh_vars) / (rh_betas + rh_lambdas) + plot_layout(guides="collect") + plot_annotation(tag_levels = 'A')

We observe that $\widehat{R}$-values of the chain means generally decreases as a function of the number of iterations (see \@ref(fig:rh)A). An exception to this observation is the initial increase when $3 \leq it \leq 5$ [interpret?? due to initialization or is there really more mixing initially??]. After the first couple of iterations, the mixing between chain means generally improves until $it \geq 30$ to $40$. There is no apparent differentiation between the missingness conditions.

The mixing between chain variances mimics the mixing between chain means almost perfectly (see \@ref(fig:rh)B). Irrespective of the missingness condition, the $\widehat{R}$-values taper off around $it \geq 30$.

With the regression estimate as parameter we observe very similar $\widehat{R}$-values than with chain means and chain variances (see \@ref(fig:rh)C). We do, however, see some differences between missingness conditions. Conditions where 75% of the cases are incomplete show more extreme non-mixing. The overall trend remains the same: about 30 iterations are required before mixing stops improving substantially.

$\widehat{R}$-values of the eigenvalues show a trend similar to the chain means and chain variances (see \@ref(fig:rh)D). [add something about conditions??]

@@

## Performance Measures

@@

# Discussion

@@

<!-- # Check if heuristic for AC works -->

@@

# Case Study
